---
title: "SexDifferencesNSCLC"
author: "Rowan Barker-Clarke"
date: "8/10/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(dplyr)
library(ggplot2)
library(gridExtra)
```

## Introduction to non-small cell lung cancer

The 5-year relative survival rate for NSCLC in women in the United States is 33%. The 5-year relative survival rate for men is 23%. For people with localized NSCLC, which means the cancer has not spread outside the lung, the overall 5-year relative survival rate is 65%.

The data we will look at is from the Lung1 cohort from the cancer genome atlas. https://wiki.cancerimagingarchive.net/display/Public/NSCLC-Radiomics

```{r read_data}
df_nsclc<-read.csv(file = "Clinical_Lung1.csv")
```

## View data

```{r view_table}
head(df_nsclc)
```

## Scatter Plot: Age at Diagnosis, Survival Time, Gender
```{r scatter, echo=FALSE}
plot_1 <- ggplot2::ggplot(aes(x=age, y=Survival.time/365, fill=gender, color=gender), data =df_nsclc %>% dplyr::filter(!is.na(Overall.Stage))) + 
  geom_point() +
  geom_smooth(method = "lm") +
  ylab("Survival (years)") +
  xlab("Age at diagnosis")

plot_1
```

## Boxplots: Age, Histology, Gender

```{r plot_data, echo=FALSE}

ggplot2::ggplot(aes(x=Overall.Stage, y=age, fill=gender, c=gender), data =df_nsclc %>% dplyr::filter(!is.na(Overall.Stage)))+ geom_boxplot() +geom_point(position = position_jitterdodge(jitter.width = 0.1))
```

```{r}
plot_2 <- ggplot2::ggplot(aes(y=Survival.time/365, x=Overall.Stage, fill=gender, c=gender), data =df_nsclc %>% dplyr::filter(!is.na(Overall.Stage))) %>% + 
  geom_boxplot(notch=TRUE) +
  geom_point(position = position_jitterdodge(jitter.width = 0.1)) + 
  ylab("Survival (years)") +
  theme(legend.position="none")
plot_2

```


```{r}
plot_3 <- ggplot2::ggplot(aes(y=Survival.trunc/365, x=Overall.Stage, fill=gender, c=gender), data =df_nsclc %>% dplyr::filter(!is.na(Overall.Stage))) %>% + 
  geom_boxplot(notch=TRUE) +
  geom_point(position = position_jitterdodge(jitter.width = 0.1)) + 
  ylab("Cancer specific survival (years)") + 
  theme(legend.position="none")
plot_3 

```

## Composite plot

```{r}
grid.arrange(plot_1, NULL, plot_2,plot_3, nrow=2, ncol=2)
```



## Extra plots



```{r}
ggplot2::ggplot(aes(y=Survival.time, x=as.factor(Clinical.N.Stage), fill=gender, c=gender), data =df_nsclc %>% dplyr::filter(!is.na(Overall.Stage))) %>% + geom_boxplot(notch=TRUE) +geom_point(position = position_jitterdodge(jitter.width = 0.1)) + ylab("Survival Time (days)")
```


```{r modify}
# truncated (cancer-specific) survival only
df_nsclc$Survival.trunc <- ifelse(df_nsclc$Survival.time>1800, NA, df_nsclc$Survival.time)

ggplot2::ggplot(aes(y=Survival.trunc/365, x=Overall.Stage, fill=gender, c=gender), data =df_nsclc %>% dplyr::filter(!is.na(Overall.Stage))) %>% + geom_boxplot(notch=TRUE) +geom_point(position = position_jitterdodge(jitter.width = 0.1)) + ylab("Survival Time (years)")

```


```{r}
# Age at death
df_nsclc$age.at.death <- (df_nsclc$age)*365+df_nsclc$Survival.time

ggplot2::ggplot(aes(x=as.factor(clinical.T.Stage), y=age.at.death/365, fill=gender, c=gender), data =df_nsclc %>% dplyr::filter(!is.na(clinical.T.Stage)))+ geom_boxplot() +geom_point(position = position_jitterdodge(jitter.width = 0.1))


```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
